---
title: "Lab04"
author: "Yating Zeng"
date: '2022-09-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
library(tidyverse)
library(data.table)
library(leaflet)
```


## Step 1. Read in the data
```{r 1.Read the data}
if (!file.exists("../lab03/met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", 
                method="libcurl", timeout = 60)
}

library(R.utils)
met <- data.table::fread("../lab03/met_all.gz")
```

## Step2. Prepare the data

Remove temperatures less than -17C

```{r remove temp<-17}
met <-met[temp >-17][elev == 9999.0, elev :=NA]
```


Make sure there are no missing data in the key variables coded as 9999, 999, etc

```{r remove-low-temp}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

# check no 9999s in other important variables


Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).

```{r create-ymd, cache = TRUE}
met <- met[ , ymd := as.Date(paste(year, month, day, sep = "-"))]
```


Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <- met[week(ymd) == 31]
```


Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
```{r check-max}
met[, .(
  temp      = max(temp,na.rm=T),
  rh        = max(rh,na.rm=T), 
  wind.sp   = max(wind.sp,na.rm=T), 
  vis.dist  = max(vis.dist,na.rm=T), 
  dew.point = max(dew.point,na.rm=T), 
  lat       = max(lat,na.rm=T), 
  lon       = max(lon,na.rm=T), 
  elev      = max(elev,na.rm=T)
)]
```

No more 9999s in our datasetï¼

```{r check-avg}
met_avg <- met[, .(
  temp      = mean(temp, na.rm=T),
  rh        = mean(rh, na.rm=T),
  wind.sp   = mean(wind.sp, na.rm=T),
  vis.dist  = mean(vis.dist, na.rm=T),
  dew.point = mean(dew.point, na.rm=T),
  lat       = mean(lat, na.rm=T),
  lon       = mean(lon, na.rm=T),
  elev      = mean(elev, na.rm=T)
), by = "USAFID"]
```


Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees

```{r}
met_avg[, region := fifelse(lon > -98 & lat >  39.71, "NE",
                    fifelse(lon < -98 & lat >  39.71, "NW",
                    fifelse(lon < -98 & lat <= 39.71, "SW", "SE")))
    ]
table(met_avg$region)

#check the missing data
#sum(is.na(met$region))
```


Create a categorical variable for elevation as in the lecture slides

```{r}
met_avg[, elev_cat := fifelse(elev >252, "high", "low")]
```


## Step 3. Use geom_violin to examine the wind speed and dew point temperature by region. Make Violin plots of dew point temp by region
```{r}
met_avg[!is.na(region)] %>%
  ggplot() +
  geom_violin(mapping = aes(x = 1, y = dew.point, color=region, fill=region)) +
  facet_wrap(~ region, nrow = 1)
```

The highest dew point temperatures are reported in the SE.

```{r}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = wind.sp,  color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 2)
```
The highest wind speed is reported in the NE.

## Step 4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region
Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph

```{r scatterplot-dewpoint-wind.sp}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp, y = dew.point)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 2)
```


## Step 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region

## Step 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x = region, y = dew.point)) +
  stat_summary(fun.data = mean_sdl, geom = "errorbar")
```

can show with means or with error bars but want to show together



#met[region=="SW", mean(dew.point,na.rm=T)]
#met[region=="NE", mean(dew.point,na.rm=T)]
#met[region=="NE", min(dew.point,na.rm)]
#met[region=="NE", sd(dew.point,na.rm)]


## Step 7. Make a map showing the spatial trend in relative h in the US

```{r}
# Generating a color palette
temp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
temp.pal
```

Use addMarkers to include the top 10 places in relative h (hint: this will be useful rank(-rh) <= 10)

```{r}
# Generating a color palette
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal
```

Use addMarkers to include the top 10 places in relative h (hint: this will be useful rank(-rh) <= 10)

```{r}
top10rh <- met_avg[ rank(-rh) <= 10]
```

```{r}
#met_avg[ order(-rh)][1:10]
```

Here we can use top 10th data subset to show just the weather stations that registered the top 10 relative humidity values in the 1st week of August.

```{r}
rhmap <- leaflet(met_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humid.', opacity=1)
rhmap
```
