---
title: "Lab05"
author: "Yating Zeng"
date: '2022-09-21'
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
library(tidyverse)
library(data.table)
library(dtplyr)
```


## Step 1. Read in the data
First download and then read in with data.table:fread()
```{r 1.Read the data}
if (!file.exists("../lab03/met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", 
                method="libcurl", timeout = 60)
}

library(R.utils)
met <- data.table::fread("../lab03/met_all.gz")
```

Remove temperatures less than -17C and change elev 9999 to missing value code.

```{r remove temp<-17}
met <-met[temp >-17][elev == 9999.0, elev :=NA]
```

Make sure there are no missing data in the key variables coded as 9999, 999, etc

```{r remove-low-temp}
met[met$elev==9999.0] <- NA
summary(met$elev)
# check no 9999s in other important variables
```


Read in the station data
```{r}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

Merge met data with stations
```{r}
met <-
  merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
nrow(met)
```

#Question 1: Representative station for the US

Compute mean temperature, wind speed, and atmospheric pressure fr each weather station, and pick the weather

```{r}
station_average <-
  met[ , .(
    temp      = mean(temp, na.rm=T),
    wind.sp   = mean(wind.sp, na.rm=T),
    atm.press = mean(atm.press, na.rm=T)
  ), by = USAFID]
```

The above computes the mean by weather station
Now let's compute the median calue for each variable.

```{r}
stmeds <- station_average[ , .(
  temp50     = median(temp, na.rm=T),
  windsp50   = median(wind.sp, na.rm=T),
  atmpress50 = median(atm.press,na.rm=T)
)]
stmeds
```

A helpful function we might want to use 'which.min()'.

```{r}
station_average[ ,
                 temp_dist50 := abs(temp - stmeds$temp50)][order(temp_dist50)]
head(station_average)
```


```{r}
station_average[ which.min(temp_dist50)]
```
It matches the result above.












