---
title: "Lab 05"
author: "Yating Zeng"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
library(dtplyr)
```


## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r read-data, cache=TRUE}
if (!file.exists("../lab03/met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}

met <- data.table::fread("../lab03/met_all.gz")
```



## 2. Prepare the data

Remove temperatures less than -17C and change elev 9999 to missing value code.
```{r remove-low-temp}
met <- met[met$temp > -17][elev == 9999.0, elev:= NA]
```


## Read in the station data

```{r stations, cache = TRUE}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

Merge met data with stations.

```{r merge_datatables}
met <- 
  merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )

nrow(met)
```

# Question 1: Representative station for the US

Compute mean temperature, wind speed and atmospheric pressure for each weather station, and pick the weather station with the average value closest to the median for the US.

```{r}
station_averages <- 
    met[ , .(
      temp      = mean(temp, na.rm = T), 
      wind.sp   = mean(wind.sp, na.rm = T),
      atm.press = mean(atm.press, na.rm = T)
    ), by = USAFID]
```

The above computes the mean by weather station.
Now let's compute the median value for each variable.

```{r}
stmeds <- station_averages[, .(
      temp50      = median(temp, na.rm = T), 
      wind.sp50   = median(wind.sp, na.rm = T),
      atm.press50 = median(atm.press, na.rm = T)
)]
stmeds
```

A helpful function we might want to use 'which.min()'

```{r}
station_averages[ , 
                  temp_dist50 := abs(temp- stmeds$temp50)][order(temp_dist50)]
```

Let's use which.min

```{r}
station_averages[ which.min(temp_dist50)]
```

It matches the result above.

# Question 2: Representative station per state
Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.

```{r} 
station_averages <- 
    met[ , .(
      temp      = mean(temp, na.rm = T), 
      wind.sp   = mean(wind.sp, na.rm = T),
      atm.press = mean(atm.press, na.rm = T)
    ), by = .(USAFID, STATE)]
head(station_averages)
```

```{r}
statemeds <- station_averages[, .(
      temp50      = median(temp, na.rm = T), 
      wind.sp50   = median(wind.sp, na.rm = T),
      atm.press50 = median(atm.press, na.rm = T)
), by = STATE]
statemeds
```

```{r}
station_averages <- merge(
  x= station_averages,
  y= statemeds,
  by.x = "STATE",
  by.y = "STATE",
  all.x= TRUE,
  all.y = FALSE
)


```


```{r}
station_averages[, temp_dist_state50 := temp-temp50]

station_averages[,windsp_dist_state50 := wind.sp - wind.sp50]

station_averages
```

```{r}
station_averages[, eucdist := temp_dist_state50^2 + windsp_dist_state50^2]
```

```{r}
repstation <- station_averages[ , .(
              eucdist = min(eucdist, na.rm=T)) 
              , by=STATE]
```

```{r}
merge(
  x= station_averages, 
  y= repstation, 
  by.x = c("eucdist", "STATE"),
  by.y = c("eucdist", "STATE"),
  all.x = FALSE, 
  all.y = TRUE
) 
```

