---
title: "Lab 03"
author: "Yating Zeng"
date: "9/7/2022"
output: github_document
always_allow_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1. Read in the data

```{r 1.Read the data}
if (!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", 
                method="libcurl", timeout = 60)
}

library(R.utils)
met <- data.table::fread("met_all.gz")
```

## Step 2. Check the dimensions, headers, footers. How many columns, rows are there?

```{r 2.Check the dimensions, headers, footers}
dim(met)
head(met)
tail(met)
```

## Step 3. Take a look at the variables.

```{r 3.Take a look at the variables}
str(met)
summary(met[,8:12])
```

## Step 4. Take a closer look at the key variables.

```{r 4.Take a closer look at the key variables}
table(met$year)
```

```{r}
table(met$day)
```

```{r}
table(met$hour)
```

```{r}
summary(met$temp)
```
```{r}
summary(met$elev)
```
```{r}
summary(met$wind.sp)
```
OR by alternative codes
```{r}
met[, summary(elev)]
summary(met$temp)
summary(met$wind.sp)
```
Check the misssing value
```{r checking the missing value}
mean(is.na(met$temp))
```

Replace elevations with 9999 as NA.

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

The weather station with the highest elevation is at `r max(met$elev, na.rm=TRUE)` meters

Remove the observation with temperature no more than -40C

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```
There is a -17.2C temperature reading that seems suspicious.

```{r}
met <- met[temp>-15][order(temp)]
```
We removed temperatures colder than -15C. The new dataset has minimum temp -3C which is reasonable.

## Step 5. Check the data against an external data source.

## Step 6. Calculate summary statistics

```{r}
met[elev==max(elev,na.rm=TRUE),summary(wind.sp)]
met[elev==max(elev,na.rm=TRUE),summary(temp)]
```


```{r}
met[elev==max(elev,na.rm=TRUE),.(
  temp_wind    = cor(temp,  wind.sp, use = "complete"),
  temp_hour    = cor(temp,   hour,   use = "complete"),
  wind.sp_day  = cor(wind.sp, day,   use = "complete"),
  wind.sp_hour = cor(wind.sp, hour,  use = "complete")
)]
```

## Step 7. Exploratory graphs

```{r}
hist(met$elev, breaks=100)
```

```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp)
```

```{r}
if (!require(leaflet)){
  install.packages("leaflet")
  library(leaflet)  
}

elev <- met[elev==max(elev)]
```
```{R the methods to show the map from Professor}
#install.packages("webshot")
#webshot::install_phantomjs()

#library(leaflet)

#library(webshot)
```

```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
#mapview::mapshot(leaflet(elev), file ="/Users/adbb/pm566-allprojects/PM566_Fall_Lab/Lab03/README_files/map.png")
```

```{r}
if (!require(lubridate)){
  install.packages("lubridate")
  library(lubridate)  
}
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```